---
title: "CBxResponsbility_Preprocessing"
author: "Carson Koffler"
format: html
editor: visual
---

## Study Background

### Quick Abstract

### Design

### Power Analysis

I will be relying on the effect size past experimenters received from using Cyberball to manipulate envy. We will likely have a bigger sample than what this requires.

> "Participants in the high-envy condition (M = 3.22, SD = 1.7) endorsed more state envy towards Standard Player 2 than did participants in the low-envy condition (M = 2.1, SD = 1.42; β = 0.33, b = 1.10, SE = 0.15, p \< .001)."

```{r pwr_t_test}
library(pwr)

z <- (pwr.t.test(n = NULL, d=.3, sig.level = 0.05, power = .80, 
           type = "two.sample", alternative = "two.sided"))

x <- (pwr.t.test(n = NULL, d=.3, sig.level = 0.05, power = .9, 
           type = "two.sample", alternative = "two.sided"))

y <- (pwr.t.test(n = NULL, d=.3, sig.level = 0.05, power = .95, 
           type = "two.sample", alternative = "two.sided"))
z
x
y

cat("\n\nAssuming d = .3 and power = .80, we need a sample of", z$n *4)
cat("\nAssuming d = .3 and power = .90, we need a sample of", x$n *4)
cat("\nAssuming d = .3 and power = .95, we need a sample of", y$n *4)


```

### Hypotheses

**Hypothesis 1:** If envy is a function of inequity, then the inequity manipulation will have an effect on episodic envy, and the blame manipulation will not.

**Hypothesis 2:** If anger is a function of blame, then the blame manipulation will have an effect on episodic anger, and the inequity manipulation will not.

**Hypothesis 3:** If money burning is a function of inequity aversion, then the inequity manipulation will have an effect on money burning, and the blame manipulation will not.

## Results

### Data Preparation

[Carson, see if you can add a variable for the trial time for the slider task specifically. I am curious if people who decided *not* to burn spent longer making that decision...]{style="color:red"}

-   Money burning is Trial 42 –\> Time elapsed at Trial 42 - Trial 41 should work?

```{r opening_file}
#| echo: false
#| include: false

#### Load Relevant Libraries and Functions
library(jsonlite)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)
library(kableExtra)
library(ggplot2)

#### Import data
# Set working directory
path <- "/Users/carsonkoffler/Documents/Envy/Cyberball/CyberballXResponsibility/Data/PilotB"
setwd(path)

files <- list.files(path = path, pattern = "\\.csv$", full.names = T) #list all files folder

length(files) #number of files --> should match n 

#read and combine participant data into one file 
all_data <- files %>%
  map_dfr(function(file_path){
    read_csv(
      file_path,
      col_types = cols(
        prolific_id_last4 = col_character(),  # force character type
        SP2_ball_received_T = col_character(),
        SP1_ball_received_T = col_character(),
        SP1_ball_received_SP2 = col_character(),
        
        .default = col_guess()                # let all other columns be guessed
      )
    )
  })

all_data <- as_tibble(all_data)

#collect whether or not participants needed more than one try on quiz 
quiz_attempts <- all_data %>%
  filter(task == "comprehension_quiz") %>%
  arrange(subject_id, trial_index)

quiz_summary <- quiz_attempts %>%
  group_by(subject_id) %>%
  summarise(
    n_attempts = n(),
    passed_first_try = first(correct),
    ever_failed = any(correct == FALSE),
    .groups = "drop"
  )

mb_time <- all_data %>%
  filter(trial_index == c(41,42)) %>%
  arrange(subject_id, trial_index)

#summarize participant data 
participants <- all_data %>%
  group_by(subject_id) %>%
  summarise(success = first(success), #complete = T/F
            experiment_data = first(experiment_date), #date study done
            exp_condition = first(condition), #condition: 1,2,3,4
            total_time = max(time_elapsed, na.rm = T), # total time to complete study
            sp1_points = first(sp1Points), 
            sp2_points = first(sp2Points),
            sp1_dollars = first(sp1Dollars),
            sp2_dollars = first(sp2Dollars), 
            burn_amount = first(burn_amount),
            experiment_date = first(experiment_date),
            prolific_id = first(prolific_id_last4),  #Last 4 of prolific id
            version = first(version)) #study version 

            ## need to figure out the From: and To: variables --> this can let me track if in
            ## condition 4 if SP1 passes to SP2 
            df_throws_list <- 
              all_data %>%
              filter(!is.na(from) & !is.na(to)) %>%
              group_by(subject_id) %>%
              summarise(
                throws = list(tibble(
                  cb_throws = trial_index,
                  from  = from,
                  to    = to
                )),
                .groups = "drop"
              ) 
            
            df_throws_list %>%
              unnest(throws)
            
            df_throws_list <- 
              all_data %>%
              filter(!is.na(from) & !is.na(to)) %>%
              select(subject_id, trial_index, condition,burn_amount, from, to)

# Identify survey rows want to keep
survey_data <- all_data %>%
  dplyr::filter(str_detect(trial_type, "survey")) %>%
  dplyr::select(subject_id, trial_type, response)

# Convert JSON strings to lists and expand (1 column per Q)
survey_expanded <- survey_data %>%
  mutate(response_list = map(response, ~ {
    # Convert string JSON to R list
    if (!is.na(.x) && .x != "") {
      as.list(fromJSON(.x))
    } else {
      NA
    }
  })) %>%
  dplyr::select(-response) %>%
  unnest_wider(response_list)


# # fix data that was not named (Q0 contains psuedonym)
# columns_fixed <- survey_expanded %>%
#   group_by(subject_id) %>%
#   summarise(
#     pseudonym = Q0[!map_lgl(Q0, is.null)][1],
#   )
# 
# #join columns_fixed to survey_expanded
# survey_expanded <- survey_expanded %>%
#   left_join(columns_fixed, by = "subject_id")
# 
# #add attention check 1 that was filtered out earlier
# attention_check_1 <- all_data %>%
#   filter(str_detect(stimulus, "Who won the last round you played\\?")) %>%
#   dplyr::select(subject_id, attention_check_1=response)
# 
# #add attention check 1 to the df 
# survey_expanded <- survey_expanded %>%
#   left_join(attention_check_1, by = "subject_id")
# 
# #rename attention check 2 to fit format 
# survey_expanded <- survey_expanded %>% 
#   rename(attention_check_2 = DV5_attention_check, 
#          psuedonum = Q0)
# 
# 
# # Join expanded survey data back to main dataframe
# # If multiple survey rows per subject, we can summarize by taking the first non-NA per column
df <- survey_expanded %>%
  group_by(subject_id) %>%
  summarise(across(everything(), ~ first(na.omit(.x)), .names = "{.col}"))

#add in the quiz data 
df <- df %>%
  left_join(quiz_summary, by = "subject_id")

df <- df %>%
  left_join(participants, by = "subject_id") ## df is now completed df 

```

Quick overview of what data looks like

```{r demo_visualizations}
#| echo: false
library(ggplot2)

#visualization of demographics 

#age
df$demo_age <- as.numeric(df$demo_age)
summary(df$demo_age)
#histogram
ggplot(data = df, aes(x= demo_age))+
  geom_histogram() +
  labs(
    title = "Distribution of Participant Ages",
    x = "Age", 
    y = "Count")+
  theme_minimal()

#gender 
#plot gender 
table(df$demo_gender)
#histogram
ggplot(data = df, aes(x= demo_gender))+
  geom_bar(fill = "#39568CFF") +
  labs(
    title = "Number of Participant's by Gender", 
    x = "Gender", 
    y = "Count")+
  theme_minimal()

#total time 
#plot time taken
summary(df$total_time/60000) #time taken in minutes ms:min --> 60000:1 
#plot
ggplot(data = df, aes(x= total_time/60000))+
  geom_density(fill = "#39568CFF") +
  labs(
    title = "Amount of Time Taken to Complete Survey",
    x = "Time to Complete Survey (minutes)", 
    y = "Count")+
  theme_minimal()


#conditions
#plot conditions 
table(df$exp_condition)

ggplot(data = df, aes(x= factor(exp_condition)))+
  geom_bar(fill = "#39568CFF") +
  labs(
    title = "Number of Participants in the Game Conditions",
    x = "Random Conditon", 
    y = "Count")+
  theme_minimal()

#attention checks 
data_complete <- df %>%
  filter(success == "TRUE")

#pass the quiz?
table(df$n_attempts)
## this will be a good indicator of how difficult it is for people to understand quiz
ggplot(data = df, aes(x= n_attempts))+
  geom_bar() +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -0.3) +
  labs(
    title = "Number of Attempts Needed to Pass Instruction Quiz", 
    x = "Number of Attempts", 
    y = "Count")+
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

#are different conditions struggling more? 
ggplot(df, aes(x = n_attempts)) +
  geom_bar(fill = "steelblue") +
  facet_wrap(~ exp_condition) +
  labs(
    title = "Number of Attempts by Experimental Condition",
    x = "Number of Attempts",
    y = "Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))


#excluded if more than 1 attention check is failed (need to pass more than 2)
#create pass/fail for attention_check_1
data_complete <- data_complete %>%
  mutate(
    comprehension_check = case_when(
      # equitable T, received ball 8
      (exp_condition == 1 & SP1_ball_received_T == 8) ~ "pass",
      #inequitable T, received ball 4
      (exp_condition == 2 & SP1_ball_received_T == 4) ~ "pass",
      #inequitable T, received ball 4
      (exp_condition == 3 & SP1_ball_received_T == 4) ~ "pass",
      #inequitable T, received ball 4
      (exp_condition == 4 & SP1_ball_received_T == 4) ~ "pass",
      TRUE ~ "fail"
    )
  )



```

#### Exclusion Criteria

```{r exclusion}
print(paste("Number of participants before any exclusions:",n_distinct(df$subject_id)))

#filter if fail 2/3 attention checks 
df_exclusion <- df %>%
  mutate(
    fail1 = dv3a_attention_check_1 != 1, 
    fail2 = demo_attention_check_2 != "1000+", 
    fail3 = attention_check_3 <= 1
  ) %>%
  filter(fail1 + fail2 + fail3 < 2)

n_excluded_attention <- n_distinct(df$subject_id) - n_distinct(df_exclusion$subject_id)

print(paste("Excluded for failed attention checks:", n_excluded_attention))
print(paste("Remaining participants:", n_distinct(df_exclusion$subject_id)))

if (n_excluded_attention != 0) {
  df <- df_exclusion
} else {
  message("No one excluded. df is good!")
}
```

### Main Analyses

```{r composites}
#| echo: false
#reverse code 
df$dv4_trade_places_calc <- 6 - df$dv4_trade_places_rev
df$dv4_content_with_self_calc <- 6 - df$dv4_content_with_self_rev
df$dv4_happy_friends_succeed_calc <- 6 - df$dv4_happy_friends_succeed_rev
df$dv4_happy_others_succeed_calc <- 6 - df$dv4_happy_others_succeed_rev

#trait envy
df$trait_envy <- (df$dv4_angry_others_succeed +
                  df$dv4_content_with_self_calc +
                  df$dv4_happy_friends_succeed_calc +
                  df$dv4_happy_others_succeed_calc +
                  df$dv4_pain_friends_succeed +
                  df$dv4_prefer_stranger_win +
                  df$dv4_rain_on_parade +
                  df$dv4_trade_places_calc)/8


#state envy composite
df$state_envy_composite <- (df$dv3_state_envy +
                            df$dv3_state_jealous +
                            df$dv3_state_resentful +
                            df$dv3_state_bitter)/4 

df$state_envy_composite2 <- (df$dv3_state_envy +
                            df$dv3_state_jealous)/2

#state anger composite 
df$state_anger_composite <- (df$dv3_state_angry +
                            df$dv3_state_annoyed +
                            df$dv3_state_upset +
                            df$dv3_state_irritated)/4 

#negative affect composite
df$state_NA_composite <- (df$dv3_state_angry +
                          df$dv3_state_annoyed +
                          df$dv3_state_bitter +
                          df$dv3_state_contempt +
                          df$dv3_state_disappointed +
                          df$dv3_state_envy +
                          df$dv3_state_irritated +
                          df$dv3_state_jealous +
                          df$dv3_state_resentful +
                          df$dv3_state_sad +
                          df$dv3_state_shame +
                          df$dv3_state_upset) /12


#negative affect (no envy)
df$state_NA_composite_NoEnvy <- (df$dv3_state_angry +
                                 df$dv3_state_annoyed +
                          
                                 df$dv3_state_contempt +
                                 df$dv3_state_disappointed +
                                  
                                 df$dv3_state_irritated +
                                  
                                  
                                 df$dv3_state_sad +
                                 df$dv3_state_shame +
                                 df$dv3_state_upset) /8



#positive affect composite 
df$state_PA_composite <- (df$dv3_state_happy +
                          df$dv3_state_joy +
                          df$dv3_state_surprised +
                          df$dv3_state_thankful +
                          df$dv3_state_compassionate +
                          df$dv3_state_grateful)/6

#negative action items
df$negative_behaviors <- (df$dv2_avoiding +
                          df$dv2_ignore   +
                          df$dv2_insulting+
                          df$dv2_unkind_things)/4

#positive action items
df$positive_behaviors <- (df$dv2_praising +
                          df$dv2_become_friends +
                          df$dv2_complementing  +
                          df$dv2_favor +
                          df$dv2_helping +
                          df$dv2_invite_fun)/6


#dummy code decision to burn 
df$burn_decision <- case_when(
                    df$burn_amount != 0 ~ 1,
                    TRUE ~ 0)

```

#### Data Visualization

What might we want to visualize?

1.  correlation matrix
2.  condition -\> envy
3.  condition -\> anger
4.  condition -\> money burning
5.  emotion -\> money burning
6.  condition -\> TAR

```{r correlation_matrix}
library(reshape2)
library(ggplot2)
dv_emotion <- grep("state_", names(df), ignore.case = TRUE, value = TRUE)


emotion_df <- df[, dv_emotion]
emotion_df <- emotion_df[,!(names(emotion_df) %in% "dv2_state_emotion_DO")]

cor_mat <- cor(
  emotion_df,
  use = "pairwise.complete.obs",
  method = "pearson"
)

# melt the correlation matrix
cor_long <- melt(cor_mat)

ggplot(cor_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +  # show correlations
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red",
    midpoint = 0, limits = c(-1,1)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8),
    axis.title = element_blank()
  ) +
  labs(fill = "Correlation")


# ggplot(df, aes(x = state_envy_composite, y = state_anger_composite)) +
#   geom_point(alpha = 0.6) +
#   geom_smooth(method = "lm", se = TRUE) +
#   facet_wrap(~ exp_condition) +
#   labs(
#     title = "Relationship Between State Envy and State Anger by Condition",
#     x = "State Envy Composite (0–6)",
#     y = "State Anger Composite (0–6)"
#   ) +
#   theme_minimal() +
#   theme(plot.title = element_text(hjust = 0.5))

ggplot(df, aes(
  x = state_envy_composite,
  y = state_anger_composite,
  color = factor(exp_condition)
)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Relationship Between State Envy and State Anger Across Conditions",
    x = "State Envy Composite (0–6)",
    y = "State Anger Composite (0–6)",
    color = "Condition"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(df, aes(
  x = state_envy_composite,
  y = state_anger_composite,
  size = burn_amount, 
  color = factor(exp_condition)
)) +
  geom_point(alpha = 0.4) +
  labs(
    title = "Relationship Between State Envy and State Anger Across Conditions",
    x = "State Envy Composite (0–6)",
    y = "State Anger Composite (0–6)",
    size = "Money Sacraficed (¢)", 
    color = "Condition"
  ) +
  scale_color_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   )) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.1))



```

1.  bar plot or box plot for the conditions on envy, anger, and

```{r hypotheses_visualizations}
# state envy by condition
ggplot(data = df, aes(x = factor(exp_condition), y = state_envy_composite, 
                      fill = factor(exp_condition)))+
  geom_boxplot() +
  stat_summary(
    fun.data = function(x) {
      data.frame(
        y = max(x, na.rm = TRUE) + 0.2,
        label = paste0("n = ", sum(!is.na(x)))
      )
    },
    geom = "text",
    size = 4
  ) +
  labs(
    title = "State Envy Across Conditions",
    x = "Conditions",
    y = "Level of State Envy (0-6)", 
    fill = "Condition"
    )+
    scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   )) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5))


# state anger by condition
ggplot(data = df, aes(x = factor(exp_condition), y = state_anger_composite, 
                      fill = factor(exp_condition)))+
  geom_boxplot() +
  stat_summary(
    fun.data = function(x) {
      data.frame(
        y = max(x, na.rm = TRUE) + 0.2,
        label = paste0("n = ", sum(!is.na(x)))
      )
    },
    geom = "text",
    size = 4
  ) +
  labs(
    title = "State Anger Across Conditions",
    x = "Conditions",
    y = "Level of State Anger (0-6)", 
    fill = "Condition"
  )+
  scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   )) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5))

#summary table of condition on 3 DVS (state envy, state anger, mb)
summary_emotion_rations <- df %>%
  group_by(exp_condition) %>%
  summarise(n = n(),
            Avg_Envy = round(mean(state_envy_composite, na.rm = T),3),
            std_Envy = round(sd(state_envy_composite, na.rm = T),3),
            Avg_Anger = round(mean(state_anger_composite, na.rm = T),3),
            std_Anger = round(sd(state_anger_composite, na.rm = T),3),
            Avg_Money_Sacrificed = round(mean(burn_amount, na.rm = T),3),
            std_Money_Sacrificed = round(sd(burn_amount, na.rm = T),3),
            .groups = "drop")

summary_emotion_rations

# Visualize Money Burning
ggplot(data = df, aes(x = burn_amount)) +
geom_histogram(bins=20)

#money burning by condition
ggplot(data = df, aes(x = burn_amount, fill = factor(exp_condition))) +
  geom_bar() + 
  scale_y_continuous(
    breaks = seq(0,145,10)
  ) +
  scale_x_continuous(
    breaks = seq(0, 12, 1),
    limits = c(-1, 12.5)
  ) +
  labs(
    title = "Amount Sacrified Across Conditions",
    x = "Sacrafice Amount (¢)",
    y = "Count", 
    fill = "Condition"
    ) +
  scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   ))+
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) 

#money buring by condition -> violin plot 
ggplot(data = df, aes(x = factor(exp_condition), y = burn_amount, 
                      fill = factor(exp_condition))) +
  geom_violin() +
  labs(
    title = "Amount Sacrified Across Conditions",
    x = "Condition",
    y = "Sacrifice Amount (¢)",
    fill = "Condition"
  ) +
  scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   ))+
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5))

#money buring by condition -> boxplot 
ggplot(data = df, aes(x = factor(exp_condition), y = burn_amount, 
                      fill = factor(exp_condition))) +
  geom_boxplot() +
  labs(
    title = "Amount Sacrified Across Conditions",
    x = "Condition",
    y = "Sacrifice Amount (¢)",
    fill = "Condition"
  ) +
  scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   ))+
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5))


#money burning and envy 
ggplot(data = df, aes(y = burn_amount, x = state_envy_composite)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "The Relationship Between Envy and Money Burning",
    y = "Sacrafice Amount (¢)",
    x = "State Envy Composite (0-6)", 
    ) +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
#money burning and anger 
ggplot(data = df, aes(y = burn_amount, x = state_anger_composite)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "The Relationship Between Anger and Money Burning",
    y = "Sacrafice Amount (¢)",
    x = "State Anger Composite (0-6)", 
    ) +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

# money burning correlated with envy and anger on same plot 

# Create a long-format dataframe for overlay
df_long <- df %>%
  pivot_longer(
    cols = c(state_envy_composite, state_anger_composite),
    names_to = "emotion",
    values_to = "score"
  )

ggplot(df_long, aes(x = score, y = burn_amount, color = emotion)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(
    values = c(
      "state_envy_composite"  = "#1b9e77", 
      "state_anger_composite" = "darkred" 
    ),
    labels = c(
      "state_envy_composite"  = "Envy", 
      "state_anger_composite" = "Anger" 
    )
  )+
  labs(
    title = "Comparing Slopes of Envy vs Anger on Money Burning",
    x = "State Emotion Composite Score (0-6)",
    y = "Sacrifice Amount (¢)",
    color = "Emotion Composite"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# anger and envy emotion ratings on the same plot (boxplot)
ggplot(df_long, aes(y = score, x = emotion, fill = factor(exp_condition))) + 
  geom_boxplot(position = position_dodge(width = 0.8)) + 
  scale_x_discrete(
    labels = c(
      "state_envy_composite"  = "Envy", 
      "state_anger_composite" = "Anger"
    )) +
  labs(
    title = "State Envy Versus State Anger Across Conditions",
    x = "Emotion Composite",
    y = "Self-Report Ratings (0-6)",
    fill = "Condition"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# anger and envy emotion ratings on the same plot (violin)
ggplot(df_long, aes(y = score, x = emotion, fill = factor(exp_condition))) + 
  geom_violin() + 
  scale_x_discrete(
    labels = c(
      "state_envy_composite"  = "Envy", 
      "state_anger_composite" = "Anger"
    )) +
  labs(
    title = "State Envy Versus State Anger Across Conditions",
    x = "Emotion Composite",
    y = "Self-Report Ratings (0-6)",
    fill = "Condition"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# anger and envy emotion ratings on the same plot (bar)
ggplot(df_long, aes(x = emotion, y = score, fill = factor(exp_condition))) + 
  stat_summary(
    fun = mean,
    geom = "col",
    position = position_dodge(width = 0.8)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  scale_x_discrete(
    labels = c(
      "state_envy_composite"  = "Envy", 
      "state_anger_composite" = "Anger"
    )
  ) +
  labs(
    title = "State Envy Versus State Anger Across Conditions",
    x = "Emotion Composite",
    y = "Mean Self-Report Rating (0–6)",
    fill = "Condition"
  ) +
  scale_fill_discrete(labels = c("Equtiable Treasurer", 
                                   "Inequitable Treasurer, No Sharing Allowed", 
                                   "Inequitable Treasurer, SP2 Can but Does Not Share",
                                   "Inequitable Treasurer, SP2 Can and Does Share"
                                   )) +
  theme_minimal() +
  theme(plot.title = element_text(margin=margin(0,0,30,0)))



## bar chart for all emotions 
df_even_longer <- df %>%
  pivot_longer(
    cols = dv_emotion,
    names_to = "emotion",
    values_to = "score"
  )

ggplot(df_even_longer, aes(x = emotion, y = score, fill = factor(exp_condition))) + 
  stat_summary(
    fun = mean,
    geom = "col",
    position = position_dodge(width = 0.8)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    title = "Self Reported Emotions Across Conditions",
    x = "Emotions",
    y = "Mean Self-Report Rating (0–6)",
    fill = "Condition"
  ) +
  scale_x_discrete(
    labels = c(
      "dv3_state_envy" = "Envy", 
      "dv3_state_jealous" = "Jealous", 
      "dv3_state_bitter" = "Bitter", 
      "dv3_state_resentful" = "Resentful", 
      "dv3_state_happy" = "Happy", 
      "dv3_state_compassionate" = "Compassionate", 
      "dv3_state_surprised" = "Surprised", 
      "dv3_state_grateful" = "Grateful", 
      "dv3_state_thankful" = "Thankful", 
      "dv3_state_sad"= "Sad", 
      "dv3_state_upset" = "Upset", 
      "dv3_state_disappointed" = "Disappointed", 
      "dv3_state_annoyed" = "Annoyed", 
      "dv3_state_angry" = "Angry",
      "dv3_state_shame" = "Shame", 
      "dv3_state_contempt" = "Contempt", 
      "dv3_state_joy" = "Joy", 
      "dv3_state_irritated" = "Irritated", 
      "state_envy_composite" = "State Envy Composite", 
      "state_envy_composite2" = "State Envy Composite 2", 
      "state_anger_composite" = "State Anger Composite", 
      "state_NA_composite" = "Negative Affect Composite", 
      "state_NA_composite_NoEnvy" = "Negative Affect (excluding envy)", 
      "state_PA_composite" = "Positive Affect Composite"
    )
  )+
  theme_minimal() +
  theme(plot.title = element_text(margin=margin(0,0,30,0)), 
        axis.text.x = element_text(angle = 45, hjust = 1))


```

Best way to demo some of the OE data we captured?

```{r wordclouds}
# #| include: false
# library(ggwordcloud)
# library(tidyverse)
# library(tidytext)
# library(RColorBrewer)
# 
# #create a text df
# df_text_long <- df %>%
#   pivot_longer(
#     cols = starts_with("dv1_"),
#     names_to = "question",
#     values_to = "text"
#   )
# 
# #remove any stop words 
# df_tokens <- df_text_long %>%
#   filter(!is.na(text), text != "") %>%
#   unnest_tokens(word, text) %>%
#   anti_join(stop_words, by = "word") %>%
#   filter(!str_detect(word, "^[0-9]+$"))
# 
# #custom stop words <-- edit this list as explore the word clouds 
# custom_stop <- tibble(
#   word = c("like", "sp2", "standard", "player", "treasurer", "ball", "throw", "throwing",
#            "pass", "passed", "passes", "threw")
# )
# 
# df_tokens <- df_tokens %>%
#   anti_join(custom_stop, by = "word")
# 
# #tokenize
# word_counts <- df_tokens %>%
#   count(question, exp_condition, word, sort = TRUE)
# 
# 
# #oe 1
# word_counts %>%
#   filter(question == "dv1_oe_sp2_behavior", n >= 1) %>%   # adjust min freq
#   ggplot(aes(label = word, size = n)) +
#   geom_text_wordcloud() +
#   scale_size_area(max_size = 15) +
#   theme_minimal() +
#   labs(title = "XXX")
# 
# word_counts %>%
#   filter(question == "dv1_oe_sp2_behavior", n >= 1) %>%
#   ggplot(aes(label = word, size = n)) +
#   geom_text_wordcloud() +
#   scale_size_area(max_size = 12) +
#   facet_wrap(~ exp_condition) +
#   theme_minimal() +
#   labs(title = "XXX")

```

```{r wordclouds_2}
# #| include: false
# df_text <- df[9:15]
# df_text$exp_condition <- df$exp_condition
# write.csv(df_text, file = "df_text.csv", row.names = FALSE)
```

#### Statistical Analyses

```{r stats}
library(dplyr)
library(emmeans)
library(lm.beta)

df$contrast1 <- case_when(
                df$exp_condition == 1 ~ -1,
                df$exp_condition == 2 ~  1, 
                df$exp_condition == 3 ~  1,
                df$exp_condition == 4 ~ -1,
                TRUE ~ -99)

df$contrast2 <- case_when(
                df$exp_condition == 1 ~ -1,
                df$exp_condition == 2 ~ -1, 
                df$exp_condition == 3 ~  3,
                df$exp_condition == 4 ~ -1,
                TRUE ~ -99)



cor.test(df$contrast1, df$contrast2)

df <- df %>%
  mutate(
    exp_condition = factor(exp_condition)
  )

#create the contrasts 
C1_raw <- c( -1, 1, 1, -1 ) 
C2_raw <- c( -1, -1, 3, -1 )


n <- as.numeric(table(df$exp_condition))

# Convert contrasts to column vectors
C1 <- matrix(C1_raw, ncol = 1)
C2 <- matrix(C2_raw, ncol = 1)

# Weighted Gram-Schmidt: make C2 orthogonal to C1
proj <- sum(n * C1_raw * C2_raw) / sum(n * C1_raw^2)
C2_orth <- C2_raw - proj * C1_raw

# Numeric codes for factor levels
level_codes <- as.numeric(df$exp_condition)

# Assign contrast values to each row
df$C1 <- C1_raw[level_codes]      
df$C2 <- C2_orth[level_codes]

#checking the orthogonality 
cor.test(df$C1, df$C2)

#check the corr across the DVS 
cor(df[, c("state_envy_composite", "state_anger_composite", "burn_amount")])

library(ppcor)

# Envy ↔ Burn decision controlling for anger
pcor.test(df$state_envy_composite,
          df$burn_decision,
          df$state_anger_composite)

# Anger ↔ Burn decision controlling for envy
pcor.test(df$state_anger_composite,
          df$burn_decision,
          df$state_envy_composite)

#hypothesis 1: condition -> envy
model_envy <- lm.beta(lm(state_envy_composite ~ C1 + C2, data = df))
summary(model_envy)
##controlling for anger
summary(lm.beta(lm(state_envy_composite ~ C1 + C2 + state_anger_composite, data = df)))

#hypothesis 2: condition -> anger
model_anger <- lm.beta(lm(state_anger_composite ~ C1 + C2, data = df))
summary(model_anger)

## controlling for envy
summary(lm.beta(lm(state_anger_composite ~ C1 + C2 + state_envy_composite, data = df)))

##dig into H2 more:
ang_model <- lm(state_anger_composite ~ factor(exp_condition), data = df)
emmeans(ang_model, pairwise ~ exp_condition, adjust = "tukey")

#hypothesis 3a: condition -> money burning decision (part 1)
model_mb_decision <- (glm(burn_decision ~ C1 + C2, 
                                 family = "binomial", data =df))
summary(model_mb_decision)

##controlling for envy and anger 
summary(glm(burn_decision ~ C1 + C2 + state_envy_composite + state_anger_composite,
    family = "binomial", data = df))



#mcfadden values for the logit models
pscl::pR2(glm(burn_decision ~ C1 + C2,family = "binomial", data = df))["McFadden"]
pscl::pR2(glm(burn_decision ~ C1 + C2 + state_envy_composite + state_anger_composite,
    family = "binomial", data = df))["McFadden"]

#variable importance for the logit models
caret::varImp(glm(burn_decision ~ C1 + C2,family = "binomial", data = df))
caret::varImp(glm(burn_decision ~ C1 + C2 + state_envy_composite + state_anger_composite,
    family = "binomial", data = df))



df_burners <- df %>%
              filter(burn_amount > 0)

#hypothesis 3b: condition -> money burning amount, giving > 0 (part 2)
model_mb <- lm.beta(lm(burn_amount ~ C1 + C2, data = df_burners))
summary(model_mb)
  ##controlling for envy and anger 
summary(lm.beta(lm(burn_amount ~ C1 + C2 + state_envy_composite + state_anger_composite, data = df_burners)))



#condition -> money burning amount, including mb = 0
summary(lm.beta(lm(burn_amount ~ C1 + C2, data = df)))

summary(lm.beta(lm(burn_amount ~ C1 + C2 + state_envy_composite + state_anger_composite, data = df)))


##dig into H3b more:
model_mb_condition <- lm(burn_amount ~ factor(exp_condition), data = df_burners)
emmeans(model_mb_condition, pairwise ~ exp_condition, adjust = "tukey")

emmeans(lm(burn_amount ~ factor(exp_condition), data = df),pairwise ~ exp_condition, adjust = "tukey")


#conditon on emotions model:
emotion_lme <- lmer(score ~ exp_condition * emotion + (1 | subject_id), 
                    data = df_even_longer)

summary(emotion_lme)

pairs(emmeans(emotion_lme, ~ exp_condition | emotion), adjust = "tukey")

contrast(
  emmeans(emotion_lme, ~ exp_condition | emotion),
  method = list("2 vs 3" = c(0, 1, -1, 0)),
  adjust = "none"
)

```

```{r thought-action-readiness}
#create long df for the TAR items
df_tar_long <- df %>%
  pivot_longer(
    cols = c("dv2_praising", "dv2_insulting","dv2_avoiding","dv2_become_friends",
             "dv2_unkind_things", "dv2_helping", "dv2_complementing", "dv2_favor",
             "dv2_invite_fun","dv2_ignore"),
    names_to = "behavior",
    values_to = "score"
  )

#barplot for TAR items across conditions
ggplot(df_tar_long, aes(x = behavior, y = score, fill = factor(exp_condition))) + 
  stat_summary(
    fun = mean,
    geom = "col",
    position = position_dodge(width = 0.8)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    title = "Tought-Action-Readiness Across Conditions",
    x = "Actions towards SP2",
    y = "Mean Self-Report Rating (0–6)",
    fill = "Condition"
  ) +
  scale_x_discrete(
    labels = c(
      "dv2_praising" = "Praise SP2", 
      "dv2_insulting" = "Insult SP2", 
      "dv2_avoiding" = "Avoid SP2", 
      "dv2_become_friends" = "Become Friends With SP2", 
      "dv2_unkind_things" = "Say Unkind Things About SP2", 
      "dv2_helping" = "Help SP2", 
      "dv2_complementing" = "Complement SP2", 
      "dv2_favor" = "Do A Favor For SP2", 
      "dv2_invite_fun" = "Invite SP2 To Do Something Fun", 
      "dv2_ignore"= "Ignore SP2"
    )
  )+
  theme_minimal() +
  theme(plot.title = element_text(margin=margin(0,0,30,0)), 
        axis.text.x = element_text(angle = 45, hjust = 1))

#negative TAR items across conditions
ggplot(data = df, aes(x= factor(exp_condition), y = negative_behaviors, fill = factor(exp_condition))) + 
  geom_boxplot() + 
  labs(
    title = "Negative Action Items Across Experimental Conditions", 
    x = "Condition", 
    y = "Negative Action Items", 
    fill = "Conditions"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust=.5))

#positive TAR items across conditions
ggplot(data = df, aes(x= factor(exp_condition), y = positive_behaviors, fill = factor(exp_condition))) + 
  geom_boxplot() +
  labs(
    title = "Positive Action Items Across Experimental Conditions", 
    x = "Condition", 
    y = "Positive Action Items", 
    fill = "Conditions"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5))

#MANOVA for condition on tar items 
tar_manova <- manova(cbind(dv2_praising, dv2_insulting, dv2_avoiding, dv2_become_friends,
             dv2_unkind_things, dv2_helping, dv2_complementing, 
             dv2_favor, dv2_invite_fun, dv2_ignore) ~ factor(exp_condition),
       data =df)

summary(tar_manova, test = "Wilks")

summary.aov(tar_manova)

#linear mixed effects -- using subject id as a random effect (behavior corr since done by the same person)
library(lme4)
tar_model <- lm(score ~ exp_condition * behavior + (1 | subject_id),
                  data = df_tar_long)

summary(tar_model)

lm(behavior ~ factor(exp_conditino), df = df_tar_long)

emmeans(tar_model, ~exp_condition | behavior, adjust = "tukey")

pairs(emmeans(tar_model, ~ exp_condition | behavior), adjust = "tukey")

#lm for contrasts on negative behaviors
summary(lm.beta(lm(negative_behaviors ~ C1 + C2, data = df)))
    # controlling for envy and anger 
summary(lm.beta(lm(negative_behaviors ~ C1 + C2 + state_envy_composite +
                                        state_anger_composite, data = df)))

#marginal means for the negative TAR items by condition
emmeans(lm(negative_behaviors ~ factor(exp_condition), data = df), pairwise ~ exp_condition, adjust = "tukey")

#lm for contrasts on positive behaviors
summary(lm.beta(lm(positive_behaviors ~ C1 + C2, data = df)))
    #controlling for envy and anger 
summary(lm.beta(lm(positive_behaviors ~ C1 + C2 + state_envy_composite +
                                        state_anger_composite, data = df)))

#marginal means for the negative TAR items by condition
emmeans(lm(positive_behaviors ~ factor(exp_condition), data = df), pairwise ~ exp_condition, adjust = "tukey")

#reverse score the postive tar items
df$dv2_become_friends_r <- 6 - df$dv2_become_friends
df$dv2_praising_r <- 6 - df$dv2_praising
df$dv2_complementing_r <- 6 - df$dv2_complementing
df$dv2_favor_r <- 6 - df$dv2_favor
df$dv2_helping_r <- 6 - df$dv2_helping
df$dv2_invite_fun_r <- 6 - df$dv2_invite_fun

#make a single tar composite 
df$tar_composite <- (df$dv2_avoiding +
                     df$dv2_become_friends_r +
                     df$dv2_complementing_r  +
                     df$dv2_favor_r +
                     df$dv2_helping_r +
                     df$dv2_ignore +
                     df$dv2_insulting +
                     df$dv2_invite_fun_r +
                     df$dv2_praising_r +
                     df$dv2_unkind_things)/10

#condition on the tar composite 
summary(lm.beta(lm(tar_composite ~ C1 + C2, data = df)))
    #controlling for envy and anger 
summary(lm.beta(lm(tar_composite ~ C1 + C2 + 
                                   state_envy_composite + state_anger_composite, 
                                   data = df)))

ggplot(data = df, aes (x = factor(exp_condition), y = tar_composite, 
                       fill = factor(exp_condition))) +
         geom_boxplot() + 
         labs(
           title = "Mean TAR score across conditions", 
           fill = "Condition", 
           x = "Condition",
           y = "Mean TAR Score (higher = more negative)"
         ) +
         theme_minimal() +
         theme(plot.title = element_text(hjust = .5))

#reassessment of dv correlation coefficients 
cor(df[, c("state_envy_composite", "state_anger_composite", "burn_amount",
           "tar_composite", "negative_behaviors", "positive_behaviors")])

library(Hmisc)
dv_corMat <- rcorr(as.matrix(df[, c("state_envy_composite", "state_anger_composite",
                                    "burn_amount","tar_composite",
                                    "negative_behaviors", "positive_behaviors")]))

dv_corMat$P

cor(df[,c("C1", "C2", "tar_composite", "negative_behaviors", "positive_behaviors")])

pcor.test(df$C1, 
          df$negative_behaviors, 
          df$C2)

pcor.test(df$C2, 
          df$negative_behaviors, 
          df$C1)

```

```{r efa}
library(psych)
fa.parallel(emotion_df, fm="pa", fa="fa") #3 factors above random
print(fa(emotion_df, nfactors=3, fm="pa", rotate="oblimin"))

state_envy_items <- c("dv3_state_envy", "dv3_state_jealous", "dv3_state_bitter", "dv3_state_resentful")

state_anger_items <- c("dv3_state_angry", "dv3_state_upset", "dv3_state_irritated", "dv3_state_annoyed")

omega(df[,state_envy_items])
omega(df[,state_anger_items])




```

```{r effects_coding}
#let's try to do effects coding instead of contrasts coding and do by condition instead of by contrast 

contrasts(df$exp_condition) <- contr.sum

summary(lm.beta(lm(state_NA_composite ~ exp_condition, data = df)))

summary(lm.beta(lm(state_envy_composite ~ exp_condition, data = df)))

summary(lm.beta(lm(state_anger_composite ~ exp_condition, data = df)))

```

```{r network}
##network analysis 
library(qgraph)
library(bootnet)

df_BxXEmo <- df[,17:44] ## grab action tendencies and emotions

corMat <- cor_auto(df_BxXEmo)

names(df_BxXEmo)
#preprocess node names for simplicity 
names <- c("1","2","3","4","5","6","7","8","9","10",
           "11","12","13","14","15","16","17","18","19","20",
           "21","22","23","24","25","26","27","28")

groups <- list("Behavior" = 1:10, 
               "Emotion"  = 11:28)

Graph_pcor <- qgraph( corMat, graph = "pcor", layout = "spring",
                      tuning = 0.25,sampleSize = nrow(df_BxXEmo), 
                      legend.cex = 0.35, vsize = 5,esize = 10,
                      posCol = "#003399", negCol = "#FF9933",vTrans = 200, nodeNames =
                      colnames(corMat), labels = names)

Graph_pcor_red <- qgraph( corMat, graph = "glasso", layout = "spring", 
                          tuning = 0.25,
                          sampleSize = nrow(df_BxXEmo), legend.cex = 0.35, 
                          vsize = 5,esize = 10, posCol = "#003399", 
                          negCol = "#FF9933",vTrans = 200, nodeNames = colnames(corMat),
                          labels = names, groups = groups, color = c("lightblue",
                                                                     "lavender"))

centralityPlot(Graph_pcor_red, include = c("Strength", "Betweenness", "Closeness"))


library(networktools)
bridge(Graph_pcor_red, communities = groups)
plot(bridge(Graph_pcor_red, communities = groups))

```
